Cross-sectoral conservation of the Indian Ocean

### 2022-02-16 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd
Added plots & histograms for fishing cost

```{r, echo = FALSE, results = "hide", message = FALSE}
options(repos = list(CRAN="http://cran.rstudio.com/")) 
library(tidyverse)
library(sf)
library(patchwork)
library(terra)
library(stars)
library(rnaturalearth)
library(rnaturalearthdata)
library(prioritizr)
library(knitr)

@@

Input type   |  Topic                 | Reference/source (format)
:-----------:|:----------------------:|:----------------------------------:
 COST        |  Expl + Res (mining)   |  ISA (shp)
 COST        |  PMN & CFC (mining)    |  Petersen et al., 2016 (shp)
 COST        |  Shipping              |  Halpern et al., 2013 (tif)
 COST        |  Fishing               |  Jase & Tin
 CONS        |  IBAs                  |  Birdlife datazone (shp)
 CONS        |  IMMAs                 |  importantmarinemammalhabitat (shp)
 CONS        |  EBSAs                 |  CBD (shp)
 CONS        |  Benthic features      |  Blue Habitats (shp)
 CONS        |  Longhurst provinces ? |  Marine Regions (shp)
 CONS        |  Vents                 |  Beaulieu & Szanfranski (2020) (csv)

@@

+ Important Bird Areas
+ Important Marine Mammal Areas
+ EBSAs
+ Deepsea features: active vents, seamounts, plateaus?

@@

@@

Targets: *need to write chunck to get ConsFeatures$amount*

@@

Seamounts,plateaus, vents

*need to cross vents with PUs*

@@

Vents: >90%

Seamounts: >70%

Plateaus: >30%

@@

Vents <- read.csv("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/vents.csv")
VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred")
VentsActiveSF<- st_as_sf(VentsActive, coords = c("Latitude", "Longitude"), crs = st_crs(PUs))
PUxVents <- st_intersects(PUs, VentsActiveSF) %>% 
    as.logical() 
PUs <- PUs %>% 
  mutate(Vents = PUxVents) ##issue: NAs everywhere, probably because trying to cross points with polygs?
```

```{r Conservation features, echo = FALSE, message = FALSE, warning = FALSE}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)

@@

```
## Cost layer

@@

+ PMN and CFC areas (Sharma, 2011 for USD valuation)
+ Exploration & reserved areas of deep-sea mining (BUT these should be locked out rather than given a cost value)
+ Fishing effort (monetary value/index? using ready-made cost layer in Spatial Planning repo?)
+ Shipping routes (lock out? or use value of cell in Halpern et al., 2013 dataset?)

@@

*Update* possibly able to have USD value for fishing & PMN (see PMN subset)

@@

```{r CFC, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
PUs <- PUs %>% 
  mutate(CFC = PUxCFC) # Add logical vector to variable called Mine1)Polyg1
```

@@

Mining layer

@@

```{r Mining layer plot, echo = FALSE, message = FALSE, warning = FALSE}
ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = CFC, aes(color="A"), show.legend = "line") + 
  geom_sf(data = PMN, aes(color="B"), show.legend = "line") + 
  geom_sf(data = Expl_PMN, aes(color="C"), show.legend = "line") + 
  geom_sf(data = Expl_PMS, aes(color="C"), show.legend = "line") +
  geom_sf(data = Expl_CFC, aes(color="C"), show.legend = "line") +
  geom_sf(data = Res_PMN, aes(color="D"), show.legend = "line") +
  geom_sf(data = Res_CFC, aes(color="D"), show.legend = "line") +
  scale_color_manual(values = c("A" = "orange", "B" = "brown", "C" = "red", "D" = "red"), 
                     labels = c("CFC", "PMN", "Exploration areas", "Reserved areas"),
                     name = "Legend") +
  ggtitle("PMN, CFC, Exploration & Reserved areas") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

```
## Shipping

**! cannot convert to polygon data when using code below (not written as code to avoid issues when knitting doc), error message says file too heavy** *will ask on Slack*

@@

```{r Shipping plot, echo = TRUE}
Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
plot(Shipping, main = 'Shipping routes')
```

@@

@@

Fishing <- fSpatPlan_Get_Cost(PUs, cCRS) ## if issues, try fx below

FishingCost <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/Cost/Cost_Raster_Sum.grd") %>% 
    terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
    st_as_sf() %>% # Convert to sf
    st_transform(cCRS) %>% # Transform to robinson
    st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
    rename(FishingCost = layer)

```
*Issues with plotting, should be a raster with colours based on cell USD value?*

```{r Fishing plot, echo = TRUE, message = FALSE, warnings = FALSE}
FishCost <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/Cost/Cost_Raster_Sum.grd")

plot(FishCost, main = 'Fishing cost')
```

@@

```
p_ConsFeat_Fishing <- problem(PUs, features = ConsFeatures, cost_column = "Fishing") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
```

@@

>>>>

```{r, echo = FALSE, results = "hide", message = FALSE}
#Open all the packages and get extras
options(repos = list(CRAN="http://cran.rstudio.com/")) 
library(tidyverse)
library(sf)
library(patchwork)
library(terra)
library(stars)
library(rnaturalearth)
library(rnaturalearthdata)
library(prioritizr)
library(knitr)

@@

Input type   |  Topic                 | Reference/source (format)
:-----------:|:----------------------:|:----------------------------------:
 COST        |  Expl + Res (mining)   |  ISA (shp)
 COST        |  PMN & CFC (mining)    |  Petersen et al., 2016 (shp)
 COST        |  Shipping              |  Halpern et al., 2013 (tif)
 COST        |  Fishing               |  Cost layer created by Jase
 CONS        |  IBAs                  |  Birdlife datazone (shp)
 CONS        |  IMMAs                 |  importantmarinemammalhabitat (shp)
 CONS        |  EBSAs                 |  CBD (shp)
 CONS        |  Benthic features      |  Blue Habitats (shp)
 CONS        |  Longhurst provinces ? |  Marine Regions (shp)
 CONS        |  Vents                 |  Beaulieu & Szanfranski (2020) (csv)

@@

+ Important Bird Areas
+ Important Marine Mammal Areas
+ EBSAs
+ Deepsea features: active vents, seamounts, rift_valleys, plateaus? *(need to define)*

@@

Conservation targets to be defined *need to write chunck*

@@

Targets: *need to write chunck*

@@

Seamounts, plateaus, active vents
Need to find out how to cross vents with PUs

@@

Vents: >90% [.csv to .shp](https://datacarpentry.org/r-raster-vector-geospatial/10-vector-csv-to-shapefile-in-r/)

Seamounts: >70% (ex: Letessier et al., 2019, PLOS Biology)

Ridges/Plateaus: >30%

@@

Vents <- read.csv("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/vents.csv")
VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred")
VentsActiveSF<- st_as_sf(VentsActive, coords = c("Latitude", "Longitude"), crs = st_crs(PUs)) # could rewrite section using pipes
```

```
PUxVents <- st_intersects(PUs, VentsActiveSF) %>% 
    as.logical() 
PUs <- PUs %>% 
  mutate(Vents = PUxVents) ##issue: NAs everywhere, probably because trying to cross points with polygs?
```

```{r Conservation features, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)

@@

## Cost layer

@@

+ PMN and CFC areas (Sharma, 2011 for USD valuation)
+ Exploration & reserved areas of deep-sea mining (lock out because won't agree to lose investments)
+ Fishing effort (using ready-made cost layer in Spatial Planning repo)
+ Shipping routes (use value of cell in Halpern et al., 2013 dataset for relative cost)

@@

@@

```{r CFC, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
PUs <- PUs %>% 
  mutate(CFC = PUxCFC) # Add logical vector to variable
```

@@

@@

```{r Mining layer plot, echo = FALSE, message = FALSE, warning = FALSE}
ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = CFC, aes(color="A"), show.legend = "line") + 
  geom_sf(data = PMN, aes(color="B"), show.legend = "line") + 
  geom_sf(data = Expl_PMN, aes(color="C"), show.legend = "line") + 
  geom_sf(data = Expl_PMS, aes(color="C"), show.legend = "line") +
  geom_sf(data = Expl_CFC, aes(color="C"), show.legend = "line") +
  geom_sf(data = Res_PMN, aes(color="D"), show.legend = "line") +
  geom_sf(data = Res_CFC, aes(color="D"), show.legend = "line") +
  scale_color_manual(values = c("A" = "orange", "B" = "brown", "C" = "red", "D" = "red"), 
                     labels = c("CFC", "PMN", "Expl_PMN", "Expl_CFC", "Res_PMN", "Res_CFC"),
                     name = "Legend") +
  ggtitle("PMN, CFC, Exploration & Reserved areas") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

## Shipping

**! cannot convert to polygon data when using code below (not written as code to avoid issues when knitting doc), error message says file too heavy** *Need to resample, will ask on Slack*

```{r Shipping plot, echo = FALSE}
Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
plot(Shipping, main = 'Shipping routes')
```

@@

@@

Could set cut-off for cost outliers at 15,000 USD or 40,000 USD - depending on whether we want to favour variation or follow standard procedures for cut-off (see histograms and maps).

```{r Fishing cost plot, echo = TRUE, message = FALSE, warnings = FALSE}

FishingCost <- fSpatPlan_Get_Cost <- function(PUs, cCRS){

  # TODO Add some filtering ability to get the cost only for specific groups

  Cost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
    terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
    st_as_sf() %>% # Convert to sf
    st_transform(cCRS) %>% # Transform to robinson
    st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
    rename(Cost = layer)
}

hist(FishingCost$Cost) #max value >200,000

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = FishingCost, aes(color=Cost)) +
ggtitle("Fishing Cost") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

FishingCostOutliers <- filter(FishingCost, Cost >= 50000) # check out outliers, play with cut-off value

hist(FishingCostOutliers$Cost)

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = FishingCostOutliers, aes(color=Cost)) +
ggtitle("Fishing Cost Outliers (>50,000 USD)") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

FishingCostNoOutliers <- filter(FishingCost, Cost <= 50000) # remove outliers, play with cut-off value

hist(FishingCostNoOutliers$Cost) # 49256 obs. = 524 outliers removed when cut-off = 20,000) but would also make sense to set cut-off at 15,000 or 40,000...

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = FishingCostNoOutliers, aes(color=Cost)) +
ggtitle("Fishing Cost Without Outliers") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

@@

@@

```
p_ConsFeat_Fishing <- problem(PUs, features = ConsFeatures, cost_column = "FishingCost") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
```

@@

[prioritizr.net](https://prioritizr.net/reference/add_locked_out_constraints.html) says to use structure below to lock out polygons but need to look more into it

```
# create problem with added locked out constraints using spatial polygon data
locked_out <- sim_pu_polygons[sim_pu_polygons$locked_out == 1, ]
p5 <- p1 %>% add_locked_out_constraints(locked_out)
```
